### 2025-05-20 docs/article/Sections/state-of-the-art.tex
chapter 2 - clearly distinguish between Kotlin coroutines and Java virtual threads.

An alternative approach involves utilizing user-level threads, while maintaining
a blocking I/O and a synchronous programming paradigm.
However, this approach still requires a user-level I/O subsystem capable of
mitigating system-level blocking, which is crucial for the performance of
I/O-intensive applications. This technique offers a lightweight solution for
efficiently managing a larger number of concurrent sessions by minimizing
per-thread overhead.
In 2020, Karsten~\cite{karsten2020} demonstrated how this strategy supports a
synchronous programming style, effectively abstracting away the complexities
associated with managing continuations in asynchronous programming.
A similar idea has been followed in Kotlin with
\textit{coroutines}~\cite{elizarov2021coroutines}, and most recently in the Java
standard library with \textit{virtual threads} released in Java 21.
% Both Spring WebFlux and Vert.X web frameworks provide native support for Kotlin
% \textit{coroutines}, enabling developers to utilize a specific
% \texttt{Dispatcher} to switch the execution thread context to a suitable
% subsystem and reduce \textit{system-level} blocking.
However, no work has yet leveraged this mechanism to enable legacy template
engines to provide PSSR while preserving non-blocking progress.
% We will demonstrate this capability by comparing it with our proposal of
% \textit{suspendable web templates} in a benchmark using Spring WebFlux.

% Determining the appropriate size of \emph{N} for the I/O susbsystem may hinge on
% factors such as the nature of the workload and the frequency of
% \textit{Blocking} I/O operations.

>>>>

An alternative approach involves utilizing user-level threads, while maintaining
a blocking I/O and a synchronous programming paradigm.
However, this approach still requires a user-level I/O subsystem capable of
mitigating system-level blocking, which is crucial for the performance of
I/O-intensive applications. This technique offers a lightweight solution for
efficiently managing a larger number of concurrent sessions by minimizing
per-thread overhead.
In 2020, Karsten~\cite{karsten2020} demonstrated how this strategy supports a
synchronous programming style, effectively abstracting away the complexities
associated with managing continuations in asynchronous programming.

Among mainstream technologies, the Kotlin programming language introduces a new 
abstraction for managing coroutines and provides structured concurrency, which 
ensures that coroutines are \textit{scoped}, \textit{cancellable}, and 
\textit{coordinated} within a well-defined 
lifecycle~\cite{elizarov2021coroutines}.
Although this model embraces the \texttt{async}/\texttt{await} 
feature~\cite{async_await}, enabling non-blocking routines to mimic the 
structure of synchronous ones and allowing developers to reason about control 
flow sequentially, it still lacks a user-level I/O subsystem and relies on an 
explicit I/O dispatcher tied to a dedicated thread pool that frees worker 
threads for handling processing tasks.
Moreover, web templates using their own templating dialects, such as JSP, 
Thymeleaf, or Handlebars, are unable to take advantage of such constructs, as 
they typically support only a limited subset of the host library's API—most 
commonly just the \texttt{Iterable} interface.

Only Java virtual threads, introduced in
JEP~444~\footnote{\url{https://openjdk.org/jeps/444}}, adopt a similar approach
to Karsten's proposal, using user-mode threads to preserve a synchronous
programming model while still allowing blocking I/O operations without tying up
platform threads. When a virtual thread performs a blocking I/O operation, the
JVM intercepts the call and transparently parks the virtual thread, freeing the
underlying platform thread to perform other tasks. Once the I/O operation
completes, the virtual thread is rescheduled on a platform thread and resumes
execution. This mechanism is enabled by the JVM's integration with non-blocking
system calls at the OS level, allowing developers to write traditional
synchronous code while benefiting from the scalability of non-blocking I/O.
